---
title: "Data Import/Cleaning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

#### Data import of eviction data *(Naama)*


#### Data import of gentrification and population density *(Will)*

In this section we'll import:

  1. NYC gentrification data from *Governing Magazine*'s [**New York City Gentrification Maps and Data**](https://www.governing.com/gov-data/new-york-gentrification-maps-demographic-data.html)
  2. Population density data from the American Community Survey (ACS) [**5-year census estimates**](https://factfinder.census.gov/faces/nav/jsf/pages/guided_search.xhtml)

```{r}
## For the ACS data, we'll have to use a crude measure of population density we calculate ourselves - census tract population (which changes every year) divided by census tract area (which does not usually change). 

## NUMERATOR
## Importing population per census tract per year


## DENOMINATOR
## Importing area in sq. mi. for each area:
area = 
  read_csv('./data/ACS_09_5YR_G001_with_ann.csv', skip = 1) %>% 
  janitor::clean_names() %>% 
  select(id, id2, geography, area_sqmi = land_area_in_square_miles) %>% 
  mutate(name = as.character(readr::parse_number(geography)),
         geography = str_remove(geography, "Census Tract [0-9]{1,}, "),
         geography = str_remove(geography, "Census Tract [0-9]{1,}\\.[0-9]{1,}, ")) %>% 
  select(id, id2, name, geography, area_sqmi)

area %>% 
  head(10) %>% 
  knitr::kable()
```


#### Data import of racial composition and English language data *(Gloria)*
