---
title: "Data Import/Cleaning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

#### Data import of eviction data *(Naama)*


#### Data import of gentrification and population density *(Will)*

In this section we'll import:

  1. NYC gentrification data from *Governing Magazine*'s [**New York City Gentrification Maps and Data**](https://www.governing.com/gov-data/new-york-gentrification-maps-demographic-data.html)  
    a. Since the data aren't available in a format that can be analyzed directly, we'll instead import data from the *Urban Displacement Project*'s [**Mapping Displacement and Gentrification in the New York Metropolitan Area**](https://www.urbandisplacement.org/maps/ny)
  2. Population density data from the American Community Survey (ACS) [**5-year census estimates**](https://factfinder.census.gov/faces/nav/jsf/pages/guided_search.xhtml)

##### Gentrification
```{r, warning = FALSE}
gentrification =
  read_excel('./data/udp_ny_final_typology_jan_2019.xlsx') %>% 
  filter(startsWith(as.character(geoid), '36')) %>% ## just New York state (FIPS = 36, first two digits)
  rename(id2 = geoid,
         gent_status = Type_1.19) %>% 
  mutate(gent_indicator = if_else(gent_status %in% c('LI - Ongoing Gentrification', 'MHI - Advanced Gentrification'),
                                  1,
                                  0))

gentrification %>% 
  arrange(id2) %>% 
  head(10) %>% 
  knitr::kable(digits = 11)
```


##### Population density
```{r, warning = FALSE}
## For the ACS data, we'll have to use a crude measure of population density we calculate ourselves - census tract population (which changes every year) divided by census tract area (which does not usually change). 

## NUMERATOR
## Importing population per census tract per year
filenames_DP05 = 
  list.files('./data/') %>%
  paste0('./data/', .) %>% 
  as_tibble() %>% 
  filter(str_detect(value, 'DP05')) %>% 
  pull(., value) ## coerces tibble back to vector for reading by map_df()

population_data = 
  map_dfr(filenames_DP05, read_csv, .id = "input") %>% 
  janitor::clean_names() %>% 
  select(id = geo_id, id2 = geo_id2, geography = geo_display_label, total_pop = hc01_vc03, year = input) %>% 
  filter(id != "Id") %>% 
  mutate(id2 = as.numeric(id2),
         year = as.numeric(year) + 2009,
         name = as.character(readr::parse_number(geography)),
         geography = str_remove(geography, "Census Tract [0-9]{1,}, "),
         geography = str_remove(geography, "Census Tract [0-9]{1,}\\.[0-9]{1,}, "),
         total_pop = as.numeric(total_pop)) %>% 
  select(id, id2, name, geography, total_pop, year)

## DENOMINATOR
## Importing area in sq. mi. for each area:
area = 
  read_csv('./data/ACS_09_5YR_G001_with_ann.csv', skip = 1) %>% 
  janitor::clean_names() %>% 
  select(id, id2, geography, area_sqmi = land_area_in_square_miles) %>% 
  mutate(name = as.character(readr::parse_number(geography)),
         geography = str_remove(geography, "Census Tract [0-9]{1,}, "),
         geography = str_remove(geography, "Census Tract [0-9]{1,}\\.[0-9]{1,}, ")) %>% 
  select(id, id2, name, geography, area_sqmi)

## JOINING DATASETS & CALCULATING DENSITY
density_data =
  left_join(population_data, area) %>% 
  mutate(density = total_pop / area_sqmi) 

## EXAMPLE DATALINES
density_data %>% 
  head(10) %>% 
  knitr::kable()
```




#### Data import of racial composition and English language data *(Gloria)*
In this section we'll import:

  1. English language usage data from the American Community Survey (ACS), with our primary measure of interest being the percentage of residents who speak English less than "very well" [**5-year census estimates**](https://factfinder.census.gov/faces/nav/jsf/pages/guided_search.xhtml)
  2. Data on racial composition by census tract from the American Community Survey (ACS) [**5-year census estimates**](https://factfinder.census.gov/faces/nav/jsf/pages/guided_search.xhtml)
```{r, warning = FALSE}
## English language usage data- we import data from 2010-2016
## Relevant variable (the percentage of the population 5 and over that speaks English less than "very well"), changes by year- files are therefore individually imported

englang_2010 = 
  read.csv("./data/ACS_10_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2010", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc170) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc170))) %>%
  select(-hc03_vc170)

englang_2011 = 
  read.csv("./data/ACS_11_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2011", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc170) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc170))) %>%
  select(-hc03_vc170)

englang_2012 = 
  read.csv("./data/ACS_12_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2012", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc170) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc170))) %>%
  select(-hc03_vc170)

englang_2013 = 
  read.csv("./data/ACS_13_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2013", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc173) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc173))) %>%
  select(-hc03_vc173)

englang_2014 = 
  read.csv("./data/ACS_14_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2014", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc173) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc173))) %>%
  select(-hc03_vc173)

englang_2015 = 
  read.csv("./data/ACS_15_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2015", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc173) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc173))) %>%
  select(-hc03_vc173)

englang_2016 = 
  read.csv("./data/ACS_16_5YR_DP02_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2016", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc173) %>%
  filter(geo_id != "Id") %>%
  mutate(pct_eng = as.numeric(as.character(hc03_vc173))) %>%
  select(-hc03_vc173)

englang_data = 
  bind_rows(englang_2010, englang_2011, englang_2012, englang_2013, englang_2014, englang_2015, englang_2016)

englang_data %>% 
  head(10) %>% 
  knitr::kable()
```

```{r, warning = FALSE}
## Racial composition data- we import data from 2010-2016
## Racial categories are not mutually exclusive, so total percentages may sum to more than 100%
## Relevant variable (the percentage of the population 5 and over that speaks English less than "very well"), changes by year- files are therefore individually imported

race_2010 = 
  read.csv("./data/ACS_10_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2010", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc72, hc03_vc73, hc03_vc74, hc03_vc75, hc03_vc76, hc03_vc77) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc72, black = hc03_vc73, aian = hc03_vc74, asian = hc03_vc75, nhpi = hc03_vc76, other = hc03_vc77) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2011 = 
  read.csv("./data/ACS_11_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2011", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc72, hc03_vc73, hc03_vc74, hc03_vc75, hc03_vc76, hc03_vc77) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc72, black = hc03_vc73, aian = hc03_vc74, asian = hc03_vc75, nhpi = hc03_vc76, other = hc03_vc77) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2012 = 
  read.csv("./data/ACS_12_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2012", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc72, hc03_vc73, hc03_vc74, hc03_vc75, hc03_vc76, hc03_vc77) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc72, black = hc03_vc73, aian = hc03_vc74, asian = hc03_vc75, nhpi = hc03_vc76, other = hc03_vc77) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2013 = 
  read.csv("./data/ACS_13_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2013", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc78, hc03_vc79, hc03_vc80, hc03_vc81, hc03_vc82, hc03_vc83) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc78, black = hc03_vc79, aian = hc03_vc80, asian = hc03_vc81, nhpi = hc03_vc82, other = hc03_vc83) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2014 = 
  read.csv("./data/ACS_14_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2014", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc78, hc03_vc79, hc03_vc80, hc03_vc81, hc03_vc82, hc03_vc83) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc78, black = hc03_vc79, aian = hc03_vc80, asian = hc03_vc81, nhpi = hc03_vc82, other = hc03_vc83) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2015 = 
  read.csv("./data/ACS_15_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2015", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc78, hc03_vc79, hc03_vc80, hc03_vc81, hc03_vc82, hc03_vc83) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc78, black = hc03_vc79, aian = hc03_vc80, asian = hc03_vc81, nhpi = hc03_vc82, other = hc03_vc83) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_2016 = 
  read.csv("./data/ACS_16_5YR_DP05_with_ann.csv") %>%
  janitor::clean_names() %>%
  mutate(year = "2016", geo_id = as.character(geo_id), geo_id2 = as.character(geo_id2)) %>%
  select(year, geo_id, geo_id2, hc03_vc78, hc03_vc79, hc03_vc80, hc03_vc81, hc03_vc82, hc03_vc83) %>%
  filter(geo_id != "Id") %>%
  rename(white = hc03_vc78, black = hc03_vc79, aian = hc03_vc80, asian = hc03_vc81, nhpi = hc03_vc82, other = hc03_vc83) %>%
  mutate(
    white = as.numeric(as.character(white)),
    black = as.numeric(as.character(black)),
    aian = as.numeric(as.character(aian)),
    asian = as.numeric(as.character(asian)),
    nhpi = as.numeric(as.character(nhpi)),
    other = as.numeric(as.character(other))
  )

race_data = 
  bind_rows(race_2010, race_2011, race_2012, race_2013, race_2014, race_2015, race_2016)

race_data %>% 
  head(10) %>% 
  knitr::kable()

```

  
  