---
title: "Neighborhood Data"
author: "Frances Williams"
date: "11/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(plotly)
library(rgdal)
library(maps)
library(devtools)
library(leaflet)
library(maptools)
library(BAMMtools)
library(patchwork)

```

```{r}
neighborhood_df = 
  read_excel("./data/NYC_neighborhoods_by_censustract_2010.xlsx", 
             skip = 5, 
             col_names = c("borough", "county_code", "boro_code", "census_tract", "PUMA", "nta_code", "neighborhood")
             ) %>% 
#creating 11 digit FIPS code by pasting country code (36) with `county_code` and `census_tract`
  mutate(geo_id = paste0("36", county_code, census_tract)
  )

```

Merging with evictions data/cleaning, weighting, and summarizing neighborhood-level data
```{r}
eviction_data = read.csv('./data/EvictionData_NY.csv') %>% 
  janitor::clean_names() %>% 
  filter(year == 2016) %>% 
  mutate(
    geoid = as.character(geoid)) %>% 
  filter(parent_location == "Kings County, New York") %>% 
  left_join(neighborhood_df, by = c("geoid" = "geo_id"))

neighborhood_data_prep = eviction_data %>% 
  group_by(neighborhood) %>% 
  mutate(
    neighb_pop = sum(population),
    tract_weight = population/neighb_pop,
    eviction_discrepancy = eviction_filing_rate - eviction_rate,
    weighted_eviction_discrep = eviction_discrepancy*tract_weight,
    weighted_eviction_rate = eviction_rate*tract_weight,
    pct_nonwhite = 100 - pct_white,
    weighted_pct_nonwhite = pct_nonwhite*tract_weight,
    weighted_povertyrate = poverty_rate*tract_weight)

neighborhood_data = neighborhood_data_prep %>% 
  summarize(
    wt_eviction_rate = as.numeric(sum(weighted_eviction_rate)),
    wt_eviction_rate_discrep = sum(weighted_eviction_discrep),
    wt_pct_nonwhite = sum(weighted_pct_nonwhite),
    wt_povertyrate = sum(weighted_povertyrate)) %>% 
  ungroup() %>% 
  mutate(neighborhood = as.factor(neighborhood))

neighborhood_data <- within(neighborhood_data, wt_eviction_rate[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)
neighborhood_data <- within(neighborhood_data, wt_eviction_rate_discrep[neighborhood == 'park-cemetery-etc-Brooklyn'] <- NA)

```

Loading neighborhood shapefile from NYC open data
```{r}
neighborhood_map = readOGR(dsn = "mapping_files/NYC_BKneighborhoods.shp", encoding = "UTF-8")


neighborhood_map@data = left_join(neighborhood_map@data, neighborhood_data, by = c("ntaname" = "neighborhood"))
```

```{r}
neighborhood_map_crs = spTransform(neighborhood_map, CRS("+init=epsg:4326"))
```

```{r}
#pop up label
label_popup = paste0(
  "<strong>Neighborhood: </strong>",
  neighborhood_map$ntaname,
  "<br><strong>Eviction rate: </strong>",
  neighborhood_map$wt_eviction_rate
)

# get jenks natural breaks
getJenksBreaks(neighborhood_map$wt_eviction_rate, 5)

#set bins based on natural breaks
eviction_rate_bins = c(0.2, 1.0, 1.8, 2.7, 3.6)

#set color palette
eviction_palette = colorBin('Reds', bins = eviction_rate_bins, na.color = '#d9d9d9')
```

Map of Eviction Rates by Neighborhood
```{r}
leaflet::leaflet(data = neighborhood_map_crs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addPolygons(
    fillColor = ~eviction_palette(wt_eviction_rate),
    fillOpacity = 0.8,
    color = "BDBDC3",
    weight = 1,
    popup = label_popup,
    highlightOptions = highlightOptions(color = "black", weight = 2, bringToFront = TRUE)) %>% 
      addLegend('bottomleft',
                pal = eviction_palette,
                values = ~wt_eviction_rate,
                title = 'Eviction rates in Brooklyn by neighborhood',
                opacity = 1)

```

Map of Eviction rate/filing discrepancy by neighborhood
```{r}

```

Stacked bar chart of eviction rate/filing discrepancy
```{r}
bar_chart_evictions = neighborhood_data %>%
  select(neighborhood, wt_eviction_rate, wt_eviction_rate_discrep) %>%
  top_n(15, wt_eviction_rate) %>% 
  mutate(neighborhood = fct_reorder(neighborhood, wt_eviction_rate)) %>%
  pivot_longer(
    cols = starts_with("wt"),
    names_to = "rate_type",
    names_prefix = "wt_",
    values_to = "rates") %>% 
  group_by(neighborhood) %>% 
  arrange(desc(rates)) %>% 
  ungroup() %>% 
  ggplot(aes(x = neighborhood, y = rates, fill = rate_type)) +
    geom_bar(position = "stack", stat = "identity") +
    coord_flip()


#HOW TO SWTICH RATE AND RATE DISCREP ON THE BAR
bar_chart_evictions

```

Bar chart of eviction rate
```{r}
barchart_evictions = neighborhood_data %>%
  top_n(15, wt_eviction_rate) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_eviction_rate),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_eviction_rate, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() + 
    labs(y = "Eviction Rate", x = "", title = "Brooklyn Neighborhood Characteristics, 2016")
  
```

Bar chart of eviction filing/rate discrepancies
```{r}
bar_chart_eviction_discrep = neighborhood_data %>%
  top_n(15, wt_eviction_rate_discrep) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_eviction_rate_discrep),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_eviction_rate_discrep, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Eviction Filing/Rate Discrepancy", x = "")

```

Bar chart of pct nonwhite
```{r}
bar_chart_pct_nonwhite = neighborhood_data %>%
  top_n(15, wt_pct_nonwhite) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_pct_nonwhite),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_pct_nonwhite, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Percent Nonwhite", x = "")

```

Bar chart of poverty rates
```{r}
bar_chart_poverty = neighborhood_data %>%
  top_n(15, wt_povertyrate) %>% 
  mutate(
    neighborhood = fct_reorder(neighborhood, wt_povertyrate),
    highlight = ifelse(neighborhood == "Brownsville" | neighborhood == "East New York (Pennsylvania Ave)" | neighborhood == "Ocean Hill" | neighborhood == "East New York" | neighborhood == "Crown Heights North", "yes", "no")) %>% 
  ggplot(aes(x = neighborhood, y = wt_povertyrate, fill = highlight)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "red", "no" = "gray"), guide = FALSE) +
    coord_flip() +
    labs(y = "Poverty Rate", x = "")

(barchart_evictions + bar_chart_eviction_discrep) / (bar_chart_pct_nonwhite + bar_chart_poverty)

```

The bar charts above take the 10 highest neighborhood values for the categories of eviction rates, eviction filing/rate discrepancy, percent non-white, and poverty rate. Five neighborhoods appear in all four charts. 

